---
title: "Lab 03"
format: html
editor: visual
---

## Lab 03 by Catherine Le

### 1. Read in the data

Sorry, this output is messy but I wasn't sure how to hide the output. I only know that the echo false will hide the code and not the output

```{r}

library(R.utils)
library(data.table)
library(tidyverse)
library(leaflet)

download.file(
  "https://raw.githubusercontent.com/USCbiostats/data-science-data/master/02_met/met_all.gz",
  destfile = file.path("~", "Downloads", "met_all.gz"),
  method   = "libcurl",
  timeout  = 60
)

met <- data.table::fread(file.path("~", "Downloads", "met_all.gz"))
```

### **2. Check the dimensions, headers, footers.**

There are 2,317,218 observations, 30 columns, and 6 rows.

```{r}
dim(met)
# head() will display the first few rows
head(met)
# tail() will display the last few rows
tail(met)
```

### **3. Take a look at the variables.**

```{r}
# displays a summary of the met dataset
str(met)
```

### **4. A closer look at the key variables**

```{r}
table(met$year)
table(met$day)
table(met$hour)

summary(met$temp)
summary(met$elev)
summary(met$wind.sp)
```

Changing elevation that is above 9999.0 to NA (missing values).

```{r}
met[met$elev==9999.0] <- NA
summary(met$elev)
```

\-- The weather station with the highest elevation is **4113 meters** (the max value in the summary table).

Removing temperatures less than 40 degrees Celsius because these are suspicious numbers.

```{r}
met <- met[temp>-40]
met2 <- met[order(temp)]
head(met2)
```

### **5. Check data against an external data source**

Removing temperatures that are greater than -15 degrees Celsius to have a more reasonable range of temperatures for this elevation.

```{r}
met <- met[temp>-15]
met2 <- met[order(temp)]
head(met2)
```

### **6. Summary Statistics**

```{r}
# selecting weather station with max elevation
elev <- met[elev==max(elev), ]
summary(elev)
```

Correlations

```{r}
# finding the correlation between temperature and wind speed
cor(elev$temp, elev$wind.sp, use="complete")
# correlation between temp and hour
cor(elev$temp, elev$hour, use="complete")
# correlation between wind speed and day 
cor(elev$wind.sp, elev$day, use="complete")
# correlation between wind speed and hour
cor(elev$wind.sp, elev$hour, use="complete")
# correlation between temp and day
cor(elev$temp, elev$day, use="complete")
```

### **7. Exploratory Graphs**

Histograms

```{r}
# histogram of elevation data with 100 bars
hist(met$elev, breaks=100)
```

Histogram above shows data is right skewed.

```{r}
# histogram of temp data
hist(met$temp)
```

Histogram above has a normal distribution

```{r}
# histogram of wind speed
hist(met$wind.sp)
```

Histogram above shows data that is right skewed

```{r}
# geograhical map of the weather station with the highest elevation
# loading magrittr to read the piping symobls
library(magrittr)
leaflet(elev) %>%
  addProviderTiles('OpenStreetMap') %>% 
  addCircles(lat=~lat,lng=~lon, opacity=1, fillOpacity=1, radius=100)
```

```{r}
# creating a date-time variable for the x-axis
library(lubridate)
elev$date <- with(elev, ymd_h(paste(year, month, day, hour, sep= ' ')))
summary(elev$date)
```

```{r}
# ordering by increasing date
elev <- elev[order(date)]
head(elev)
# looking at the first couple of rows
```

Plotting the date-time variable we just created

```{r}
plot(elev$date, elev$temp, type='l')
```

The temperatures seem to be high around August 6th and August 20th. The temperatures were at an all-time low at August 27th

```{r}
plot(elev$date, elev$wind.sp, type='l')
```

The wind speed was at a high on August 17th and August 27th.
